\thispagestyle{plain}

\justifying


As with all things in life, all good things come to an end. This is the end of my journey at Northeastern University after six eventful years of growth, learning and immense experience with amazing people. I like to thank God for life, my parents - Dennis and Ifeoma whose prayers, emotional support and resources have brought me this far I remain blessed thank you for everything both professional, parental and spiritual. My siblings; Kachi, Ijem and Nchedo especially, Nchedo he's been an amazing brother and friend. I also want to mention Obehi, he has become family,  a confidant and brother whose guidance especially in navigating the difficult terrain of the professional world has been invaluable. My mentor Prof. Marilyn Minus has been a rock throughout my time at Northeastern. She has always encouraged me and given me the advice to keep fighting despite the challenges. I am also very happy to have met Prof. Andrew Gouldstone, who introduced me to Prof. Minus he's jovial yet has an amazing ability to convey excellent analysis and sit down to interact with just about anyone. It is something I have always loved about him and admire. Prof. Kowalski, he was someone who saw talent in me and helped me nurture it through close interactions and eventually convincing me to stay in grad school. I'll like to also mention Prof. Smyser, she is an amazing person and I hope to work with people like her. She showed me empathy in leadership and how to appreciate the people you work with even if they're direct reports. I always marvelled at the thought of how she cared for her grad students by the timely Christmas gift cards year after year. Also, Kevin McCue was someone I got to know and respect. We had close interactions and he made me confident in front of students as a TA, his pragmatic approach to problem solving really helped shape me as an engineer. 

In MIE I have never met somone as impressionable as Prof. Mehdi Abedi, his love and passion for science pushed me to the brink. Likewise his sometimes encouraging words. Prof. Metghalchi was my advisor you were kind and gracious I always recall the first time we met and how you always reminded us to use the right lingo when discussing energy and it's become a lifestyle I like to imbibe. Prof. Levendis, after we met you impressed me with your style and sense of fashion, your dedicated passion for combustion pushed me to learn about gas turbine combustion and we've become life long friends and I always cherished being able to walk into your office to discuss about combustion thank you for the memories. 

Prof. Michael Allshouse thank you so much for the exposure you gave me I never believed I would be able to code and here I am writing a thesis about code generation in chemical kinetics. Thank you for your kind words even when we meet at church you and Nina are always so welcoming. You made me love CFD despite it being the hardest class I took, you made me like your ability to debug code and interpret what the solver was doing, a very methodological approach. 

Prof. Richard West, thank you so much. This thesis is really all about your help for the last two years. I know how I met you and got frustrated not knowing what to work on yet you persisted and pushed me beyond my abilities. Because of your guidance I learned to like chemistry and kinetics again. I know deep down I did frustrate you as a student but you never gave up on me. Thank you for introducing me to RMG and the Cantera team with Dr. Ray Speth, Prof. DeCaluwe, Prof. Bryan Weber, Prof. Kyle Niemeyer as well as chemical kinetics I do hope to become as proficient at coding as you one day. To the CoMoCheng group Nate, Krishna, Emily, David, Chris, Shenghui and Chao you all contributed passively to me learning in one way or the other I hope to form the bond you all share. To the MIT RMG team especially, Matt Johnson, you'll forever be in my heart your graciousness to always teach and explain code and debug issues in RMG made this thesis possible and Max Liu and Mark Payne. Thank you for the late night responses, impromptu visits and calls. Dr. Richard Marsh at Cardiff University who sparked my interest in combustion thank you for this gift.

To my close friends who pushed me harder than they could imagine Nneoma, Emmanuel, Oseghale, Uzoma for helping me out with python and data science, Natasha and my recent but go to workout mentor, my wonder woman Chinenye you pushed me harder than I could imagine and provided the emotional support to go through it all. I remain grateful. To the podcasts from Ravi Zacharias, Steven Furtick, Pastor Jan, Pastor Sola Osunmakinde my Back Bay CG at Mosaic, those podcasts and messages lifted my spirit in times of despair thank you. 

To anyone who'll ever read this thesis, be encouraged to never despair in times of hopelessness and adversity even when all hopes seem lost, continue striving for it is through our steepest moments of uncertainty is our greatest work conceived. Thank you to all my teachers from primary school to secondary school and at Oxbridge for the investment over the years. It was an honor to have learned from you all.